---
title: "Prediction Assignment"
output: html_document
---

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. 
In this project, in order to quantify how well an activity is done, the goal is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

## Data

The training data for this project are available here:
  
  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
  
  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. 

## Installing Packages
```{r pack,warning=FALSE,message=FALSE}
IscaretInstalled <- require("caret")
if(!IscaretInstalled){
  install.packages("caret")
  library("caret")
}

IsrandomForestInstalled <- require("randomForest")
if(!IsrandomForestInstalled){
  install.packages("randomForest")
  library("randomForest")
}

IsRpartInstalled <- require("rpart")        
if(!IsRpartInstalled){
  install.packages("rpart")
  library("rpart")
}

IsRpartPlotInstalled <- require("rpart.plot")        
if(!IsRpartPlotInstalled){
  install.packages("rpart.plot")
  library("rpart.plot")
}

#Set seed for reproducability
set.seed(20000)
```

## Data Processing
```{r process}
#Load the data and delete missing values
trainUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"   
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))  
testing <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))

#Remove variables with near zero variance
training<-training[,colSums(is.na(training)) == 0]
testing <-testing[,colSums(is.na(testing)) == 0]

#Remove columns that are not predictors, which are the the seven first columns
training   <-training[,-c(1:7)]
testing <-testing[,-c(1:7)]


```

## Cross-validation
```{r cross}
#In order to get out-of-sample errors, split the training data in training (75%) and testing (25%) data) subsets:
inTrain <- createDataPartition(y=training$classe, p=0.75, list=FALSE)    
CrossTraining <- training[inTrain, ]
CrossTesting <- training[-inTrain, ]  
dim(CrossTraining)
dim(CrossTesting)
```

## Prediction Models
```{r model}
#DECISION TREE
#Fit model on CrossTraining data
fitDTree <- rpart(classe ~ ., data=CrossTraining, method="class")

#Use model to predict class in validation set (CrossTesting)
predictionDTree <- predict(fitDTree, CrossTesting, type = "class")

#Plotting the decision tree
prp(fitDTree)

#Estimate the errors of the prediction algorithm in the Decision Tree model
confusionMatrix(CrossTesting$classe, predictionDTree)

#RANDOM FOREST
#Fit model on CrossTraining data
fitRForest <- randomForest(classe ~ ., data=CrossTraining, method="class")

#Use model to predict class in validation set (CrossTesting)
predictionRForest <- predict(fitRForest, CrossTesting, type = "class")

#Estimate the errors of the prediction algorithm in the Random Forest
confusionMatrix(CrossTesting$classe, predictionRForest)
```

# Testing the model to predict 20 test cases
```{r quiz}
# Perform prediction
predictSubmission <- predict(fitRForest, testing, type="class")
predictSubmission

```